---
title: "p8105_hw1_ga2612"
author: "Lupe Antonio"
date: "9/21/2023"
output: html_document
---

# Problem 1
```{r setup, include=FALSE}
library(moderndive)
data(early_january_weather)
```

```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

```{r, message=FALSE}
library(tidyverse)
mean(pull(early_january_weather, temp))
```


#### Data Description

The `early_january_weather` dataset contains weather data for EWR (`origin`) for the first 15 days of January in 2013. The dataset consists of 358 rows/cases with 15 columns. Each row corresponds to one hour of each day of the first 15 days of January being reported. For each hour, the information reported consists of `temp` = temperature in F, `dewp` = dewpoint in F, `humid` = relative humidity, `wind_dir` = wind direction in degrees, `wind_speed` = wind speed in mph, `wind_gust` = gust speed in mph, `precip` = precipitation in inches, ` pressure` = sea level pressure in millibars, and `visib` = visibility in miles. Additionally, there is a variable (`time_hour`) that corresponds to the date and hour of the recording of information for each row/case. Based on the dataset, the average temperature for those first 15 days of January was about 39.58 degrees Fahrenheit. 



#### Scatterplot
```{r}
library(ggplot2)

ggplot(data = early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point()
```

```{r, include=FALSE}
#saving plot pdf
ggsave("early_january_weather.pdf")
```

#### Patterns description

The plot above represents the temperature (y-axis) for each hour for the first 15 days of January 2013 (x-axis). Based on the plot, we can see that as higher tempertures were reported, lower levels of humidity were reported. Additionally, we can observe that temperatures tend to flucuate between hours, but most drastic changes can be noted between days.




# Problem 2

```{r}
df2 <- tibble(
  sd_norm_samp = rnorm(n = 10, mean = 0, sd = 1),
  norm_samp_pos = sd_norm_samp > 0,
  vec_char = c("Hi", "my", "name", "is", "Lupe", "and", "I", "like", "data", "science"),
  vec_fac = factor(c("purple", "lavender", "lilac", "lavender", "lilac", "lavender", "purple", "lilac", "lavender", "lilac"))
)
```

```{r}
#mean for the random sample
mean(pull(df2, sd_norm_samp))
```

```{r}
#mean for logical vector
mean(pull(df2, norm_samp_pos))
```

Taking the mean of the normal sample and logical vector works, but it is not possible to take the mean of the character and factor variables.


```{r, eval = FALSE}
df2 <- df2 %>%
  mutate(
    norm_samp_pos = as.numeric(norm_samp_pos),
    vec_fac = as.numeric(vec_fac),
    vec_char = as.numeric(vec_char)
  )
```


When trying to apply the `as.numeric` function to the logical, factor, and character variables, the function only works with the logical and factor character. The function works with the logical variables because it is a binrary variable, so it converts it to 0's and 1's. Additionally, since the factor variable has three levels, it converts the levels into numerical variables such as 1's, 2's, and 3's. Alternatively, the function doesn't work for the character variable because there is no way to categorize character variables. Thus, it doesn't allow for its mean to be taken. 
